---
title: "EDA_Popular Game Platform Trends"
format: html
---

```{r}
# import needed libraries
library(dplyr)
library(tidyverse)
library(ggplot2)
library(wordcloud)
library(wordcloud2)
library(grid)
library(viridis)

# read in data
df <- read.csv("data/cleaned_game.csv")
```

# Popular Game Platform Trends

## Explore the changing popularity of gaming platforms over time.
Since there are so many console types, I will first choose the first ten consoles that have the most number of released games
```{r}
# extract the required attributes
df_2 <- df |> select(console,release_date)

# check if there are any NA rows
colSums(is.na(df_2))

# drop NA
df_2 <- na.omit(df_2)

# transform date 
df_2$release_date <- as.Date(df_2$release_date, format="%Y-%m-%d")
df_2$release_y <- format(df_2$release_date, "%Y")

# count the numbers based on the released dates and consoles 
console_release_count <- df_2 |> group_by(console) |> summarise(release_count = n(), .groups = "drop") |> arrange(desc(release_count))

# fetch the top ten consoles and try to show their changing pattern of the number of the released games with released year
top_10_consoles <- head(console_release_count, 10)
top_10_consoles <- df_2 |> filter(console %in% top_10_consoles$console)
trend_data <- top_10_consoles |> group_by(console, release_y) |> summarise(release_count = n(), .groups = "drop")

trend_data <- trend_data |> filter(release_y >= 1995 & release_y <= 2019)
print(trend_data)

# plot the heatmap
heatmap_data <- trend_data |> pivot_wider(names_from = console, values_from = release_count, values_fill = 0)
heatmap_data_long <- pivot_longer(heatmap_data, cols = -release_y, names_to = "console", values_to = "count")
heatmap_data_long |> ggplot(aes(x = release_y, y = console, fill = count)) +
                     geom_tile(color = "grey70")+
                     scale_fill_gradient(low = "white", high = "blue") +
                     labs(title = "Top 10 Game Consoles: Released Games Over Years (1995-2018)",
                          x = "Release Year",
                          y = "Consoles",
                          fill = "Number of Games") +
                     theme_minimal()+
                     theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )

# save the plot
ggsave("Console_release_games.png", path = "Image_for_Report/Platform", width = 14, height = 10, dpi = 300)
```

Due to the large number of console types in the dataset, which makes the resulting plot difficult to interpret, I will group all consoles into broader console systems for a more cohesive analysis.
```{r}
# extract the required attributes
df_1 <- df |> select(console,release_date)

# check if there are any NA rows
colSums(is.na(df_1))

# drop NA
df_1 <- na.omit(df_1)

# transform consoles
df_1 <- df_1 |> mutate(console_series = case_when(console %in% c("PS", "PS2", "PS3", "PS4", "PS5") ~ "PlayStation",
                                                  console %in% c("XB", "X360", "XBL", "XOne") ~ "Xbox",
                                                  console %in% c("Wii", "WiiU", "NS") ~ "Nintendo",
                                                  console %in% c("DS", "3DS", "GB", "GBA", "GBC") ~ "Nintendo Handheld",
                                                  console %in% c("PSP", "PSV") ~ "PlayStation Handheld",
                                                  console %in% c("GEN", "SAT", "SNES", "DC") ~ "Sega",
                                                  console %in% c("PC") ~ "PC",
                                                  TRUE ~ "Other"))

# transform date 
df_1$release_date <- as.Date(df_1$release_date, format="%Y-%m-%d")
df_1$release_y <- format(df_1$release_date, "%Y")

# plot the heatmap
year_count <- df_1 |> group_by(console_series, release_y) |> summarise(release_count = n(), .groups = "drop")
year_count <- year_count |> filter(release_y >= 1995 & release_y <= 2019)

heatmap_data <- year_count |> pivot_wider(names_from = console_series, values_from = release_count, values_fill = 0)
heatmap_data_long <- pivot_longer(heatmap_data, cols = -release_y, names_to = "console", values_to = "count")
heatmap_data_long |> ggplot(aes(x = release_y, y = console, fill = count)) +
                     geom_tile(color = "grey70")+
                     scale_fill_gradient(low = "white", high = "blue") +
                     labs(title = "Game Console Series: Released Games Over Years (1995-2018)",
                          x = "Release Year",
                          y = "Console Series",
                          fill = "Number of Games") +
                     theme_minimal()+
                     theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )
# save the plot
ggsave("Console_series_release_games.png", path = "Image_for_Report/Platform", width = 16, height = 13, dpi = 300)
```

Since the resulting plot is still difficult to interpret, I plan to break it down into two smaller topics:
1. Comparison of Game Releases Between Nintendo and Sony (Competitive Relationship) with the change of time
```{r}
# extract Nintendo data and Sony data
Nintendo = df_1 |> filter(console_series %in% c("Nintendo Handheld", "Nintendo"))
Sony = df_1 |> filter(console_series %in% c("PlayStation Handheld", "PlayStation"))

# count the number of released games by two brands
Nintendo$brand <- "Nintendo"
Sony$brand <- "Sony"
year_release_count_Nintendo <- Nintendo |> group_by(release_y, brand) |> summarise(release_count = n(), .groups = "drop") 
year_release_count_Sony <- Sony |> group_by(release_y, brand) |> summarise(release_count = n(), .groups = "drop") 
df_plot <- rbind(year_release_count_Nintendo, year_release_count_Sony)

# calculate the mean number of released game
df_mean <- data.frame(
        group <- c("Nintendo","Sony"),
        mean <- c(mean(year_release_count_Nintendo$release_count),mean(year_release_count_Sony$release_count))
)

# plot the result
df_plot |> ggplot(aes(x = release_y, y = release_count, color = brand, group = brand)) +
                        geom_line(size = 1) +  
                        geom_point(size = 5) +  
                        geom_hline(data = df_mean, aes(yintercept = mean, color = group), linetype = "dotted", size = 3, alpha = 0.8)+
                        labs(title = "Released Games Over Years: Sony vs Nintendo",
                             subtitle = "Number of game releases per year by each brand",
                             x = "Year",
                             y = "Number of Releases Per Year",
                             color = "Brand") +
                        theme_minimal()+
                        theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )

# save the plot
ggsave("Released Games Over Years：Sony vs Nintendo.png", path = "Image_for_Report/Platform", width = 14, height = 10, dpi = 300)
```

2. Comparison Between PC, Console, and Handheld Platforms with the change of time
```{r}
# extract data
PC = df_1 |> filter(console_series %in% c("PC"))
Console <- df_1 |> filter(console_series %in% c("Xbox", "PlayStation", "Nintendo"))
Handheld <- df_1 |> filter(console_series %in% c("PlayStation Handheld", "Nintendo Handheld"))

# count the number of released games by two brands
PC$platform <- "PC"
Console$platform <- "Console"
Handheld$platform <- "Handheld"
year_release_count_PC <- PC |> group_by(release_y, platform) |> summarise(release_count = n(), .groups = "drop")
year_release_count_Console <- Console |> group_by(release_y, platform) |> summarise(release_count = n(), .groups = "drop") 
year_release_count_Handheld <- Handheld |> group_by(release_y, platform) |> summarise(release_count = n(), .groups = "drop") 
df_plot <- rbind(year_release_count_PC, year_release_count_Console, year_release_count_Handheld)

# calculate the mean number of released game
df_mean <- data.frame(
        group <- c("PC","Console", "Handheld"),
        mean <- c(mean(year_release_count_PC$release_count),mean(year_release_count_Console$release_count), mean(year_release_count_Handheld$release_count))
)

# plot the result
df_plot |> ggplot(aes(x = release_y, y = release_count, color = platform, group = platform)) +
                        geom_line(size = 1) +  
                        geom_point(size = 5) +  
                        geom_hline(data = df_mean, aes(yintercept = mean, color = group), linetype = "dotted", size = 3, alpha = 0.8)+
                        labs(title = "Released Games Over Years: PC vs Consoles vs Handhelds",
                             subtitle = "Annual number of game releases by platform type",
                             x = "Release Year",
                             y = "Number of Releases",
                             color = "Platform") +
                        theme_minimal()+
                        theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )
  

# save the plot
ggsave("Released Games Over Years：PC vs Console vs Handheld.png", path = "Image_for_Report/Platform", width = 14, height = 10, dpi = 300)
```

I will also evaluate the platforms based on sales.
Plotting based on platforms
```{r}
# extract the required attributes
df_3 <- df |> select(console,release_date, total_sales)

# check if there are any NA rows
colSums(is.na(df_3))

# drop NA
df_3 <- na.omit(df_3)

# transform date 
df_3$release_date <- as.Date(df_3$release_date, format="%Y-%m-%d")
df_3$release_y <- format(df_3$release_date, "%Y")

# count the numbers based on the released dates and consoles 
console_sale_count <- df_3 |> group_by(console) |> summarise(release_count = n(), .groups = "drop") |> arrange(desc(release_count))


# fetch the top ten consoles and try to show their changing pattern of the number of the released games with released year
top_10_consoles <- head(console_sale_count, 10)
top_10_consoles <- df_3 |> filter(console %in% top_10_consoles$console)
trend_data <- top_10_consoles |> group_by(console, release_y) |> summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = "drop")
trend_data <- trend_data |> filter(release_y >= 1995 & release_y <= 2019)


# plot the heatmap
heatmap_data <- trend_data |> pivot_wider(names_from = console, values_from = total_sales, values_fill = 0)
heatmap_data_long <- pivot_longer(heatmap_data, cols = -release_y, names_to = "console", values_to = "count")
heatmap_data_long |> ggplot(aes(x = release_y, y = console, fill = count)) +
                     geom_tile(color = "grey70")+
                     scale_fill_gradient(low = "white", high = "blue") +
                     labs(title = "Top 10 Game Consoles: Game Sales Over Years (1995-2018)",
                          x = "Year",
                          y = "Consoles",
                          fill = "Sales (in millions)") +
                     theme_minimal()+
                     theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )
# save the plot
ggsave("Console_sales_games.png", path = "Image_for_Report/Platform", width = 14, height = 10, dpi = 300)
```

Plotting based on the types of the platforms
```{r}
# transform consoles
df_4 <- df_3 |> mutate(console_series = case_when(console %in% c("PS", "PS2", "PS3", "PS4", "PS5") ~ "PlayStation",
                                                  console %in% c("XB", "X360", "XBL", "XOne") ~ "Xbox",
                                                  console %in% c("Wii", "WiiU", "NS") ~ "Nintendo",
                                                  console %in% c("DS", "3DS", "GB", "GBA", "GBC") ~ "Nintendo Handheld",
                                                  console %in% c("PSP", "PSV") ~ "PlayStation Handheld",
                                                  console %in% c("GEN", "SAT", "SNES", "DC") ~ "Sega",
                                                  console %in% c("PC") ~ "PC",
                                                  TRUE ~ "Other"))

# plot the heatmap
year_count <- df_4 |> group_by(console_series, release_y) |> summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = "drop")
year_count <- year_count |> filter(release_y >= 1995 & release_y <= 2019)

heatmap_data <- year_count |> pivot_wider(names_from = console_series, values_from = total_sales, values_fill = 0)
heatmap_data_long <- pivot_longer(heatmap_data, cols = -release_y, names_to = "console", values_to = "count")
heatmap_data_long |> ggplot(aes(x = release_y, y = console, fill = count)) +
                     geom_tile(color = "grey70") +
                     scale_fill_gradient(low = "white", high = "blue") +
                     labs(title = "Game Console Series: Game Sales Over Years (1995-2018)",
                          x = "Year",
                          y = "Console Series",
                          fill = "Sales (in millions)") +
                     theme_minimal()+
                     theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )
# save the plot
ggsave("Console_series_sales_games.png", path = "Image_for_Report/Platform", width = 16, height = 13, dpi = 300)
```

1. Comparison of Game Sales Between Nintendo and Sony (Competitive Relationship) with the change of time
```{r}
# extract Nintendo data and Sony data
Nintendo = df_4 |> filter(console_series %in% c("Nintendo Handheld", "Nintendo"))
Sony = df_4 |> filter(console_series %in% c("PlayStation Handheld", "PlayStation"))

# count the number of released games by two brands
Nintendo$brand <- "Nintendo"
Sony$brand <- "Sony"
year_sale_Nintendo <- Nintendo |> group_by(release_y, brand) |> summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = "drop") 
year_sale_Sony <- Sony |> group_by(release_y, brand) |> summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = "drop") 
df_plot <- rbind(year_sale_Nintendo, year_sale_Sony)

# calculate the mean number of released game
df_mean <- data.frame(
        group <- c("Nintendo","Sony"),
        mean <- c(mean(year_sale_Nintendo$total_sales),mean(year_sale_Sony$total_sales))
)

# plot the result
df_plot |> ggplot(aes(x = release_y, y = total_sales, color = brand, group = brand)) +
                      geom_line(size = 1) +  
                      geom_point(size = 5) +  
                      geom_hline(data = df_mean, aes(yintercept = mean, color = group), linetype = "dotted", size = 3, alpha = 0.8) +
                      labs(title = "Game Sales Over Years: Sony vs Nintendo",
                             subtitle = "Number of game sold per year by each brand",
                             x = "Year",
                             y = "Number of Sales Per Year (in millions)",
                             color = "Brand") +
                      theme_minimal()+
                      theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )

# save the plot
ggsave("Game Sales Over Years：Sony vs Nintendo.png", path = "Image_for_Report/Platform", width = 14, height = 10, dpi = 300) 
```

2. Comparison of Game Sales Between PC, Console, and Handheld Platforms with the change of time
```{r}
# extract data
PC = df_4 |> filter(console_series %in% c("PC"))
Console <- df_4 |> filter(console_series %in% c("Xbox", "PlayStation", "Nintendo"))
Handheld <- df_4 |> filter(console_series %in% c("PlayStation Handheld", "Nintendo Handheld"))

# count the number of released games by two brands
PC$platform <- "PC"
Console$platform <- "Console"
Handheld$platform <- "Handheld"
year_sales_PC <- PC |> group_by(release_y, platform) |> summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = "drop")
year_sales_Console <- Console |> group_by(release_y, platform) |> summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = "drop") 
year_sales_Handheld <- Handheld |> group_by(release_y, platform) |> summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = "drop") 
df_plot <- rbind(year_sales_PC, year_sales_Console, year_sales_Handheld)

# calculate the mean number of released game
df_mean <- data.frame(
        group <- c("PC","Console", "Handheld"),
        mean <- c(mean(year_sales_PC$total_sales),mean(year_sales_Console$total_sales), mean(year_sales_Handheld$total_sales))
)

# plot the result
df_plot |> ggplot(aes(x = release_y, y = total_sales, color = platform, group = platform)) +
                        geom_line(size = 1) +  
                        geom_point(size = 5) +  
                        geom_hline(data = df_mean, aes(yintercept = mean, color = group), linetype = "dotted", size = 3, alpha = 0.8)+
                        labs(title = "Game Sales Over Years: PC vs Consoles vs Handhelds",
                             subtitle = "Annual number of game sold by platform type",
                             x = "Year",
                             y = "Number of Sales (in millions)",
                             color = "Platform") +
                        theme_minimal()+
                        theme(plot.title = element_text(color = "black",hjust = 0.5, size = 25, face = "bold"),
                           axis.title = element_text(color = "black",size = 25, face = "bold"),
                           axis.text = element_text(color = "black",size = 20, face = "bold"),
                           axis.text.x = element_text(color = "black",angle = 45, hjust = 1),
                           legend.title = element_text(color = "black", size = 20,face = "bold"),
                           legend.text = element_text(color = "black", size = 20,face = "bold"),
                           panel.background = element_rect(fill = "white"), 
                           plot.background = element_rect(fill = "white"), 
                           legend.background = element_rect(fill = "white") )

# save the plot
ggsave("Game Sales Over Years：PC vs Console vs Handheld.png", path = "Image_for_Report/Platform", width = 14, height = 10, dpi = 300)

```